---
title: "Descriptive statistics for manuscript"
author: "Tuomas Aivelo"
output: html_document
---


First we start with loading required packages


```{r, message=FALSE, warning=FALSE}
deps = c("vegan","dae","plyr","scales","lme4","nlme");

for (dep in deps){
  if (dep %in% installed.packages()[,"Package"] == FALSE){
    install.packages(as.character(dep), quiet=TRUE);
  }
  library(dep, verbose=FALSE, character.only=TRUE)
}
```

Then we combine all the data and do some housekeeping on the data

```{r}
md <- read.csv(file="interaction_metadata.csv", sep=";", dec=",")
otu.alpha <- read.table(file = "microbiota.ave-std.summary", 
                        header = T)
otu.alpha_ave <- subset(otu.alpha, method=="ave") 
md <- merge(md,otu.alpha_ave, by="group")

rownames(md) <- md$group
nsamples <- nrow(md)
o <- order(rownames(md))
md <- md[o, ]
et.count <- table(md$name)
dup.name <- names(et.count[et.count>1])

completeFun <- function(data, desiredCols) {
  completeVec <- complete.cases(data[, desiredCols])
  return(data[completeVec, ])
}
md <- completeFun(md, "cestode")

md$week[md$week=="39"]<-40
md$week[md$week=="44"]<-43
```

Let's start with richness, first with full model and then with reduced model

```{r}
anova(lme(sobs ~ pcrbatch+seqbatch+age+aggr+cond+sex+site+week+nematode+cestode+eimeria, data=md, random=~1|name))
anova(lme(sobs ~ age+aggr+cond+sex+site+week+nematode+cestode+eimeria, data=md, random=~1|name))

```

Though if we reduce too much, even Eimeria is non-significant!

```{r}
anova(lme(sobs ~ eimeria, data=md, random=~1|name))
```

Let's use the more complete model to calculate the values.

```{r}
summary(lme(sobs ~ age+aggr+cond+sex+site+week+nematode+cestode+eimeria, data=md, random=~1|name))

```

Eimeria is the only significant covariate.

And let's repeat this for diversity, as calculated by inverse Simpson index.

```{r}
anova( lme(invsimpson ~pcrbatch+seqbatch+age+aggr+cond+sex+site+week+nematode+cestode+eimeria+ecto, data=md, random=~1|name))
anova(lme(invsimpson ~ eimeria, data=md, random=~1|name)) #all <0.05!
summary(lme(invsimpson ~ age+aggr+cond+sex+site+week+nematode+cestode+eimeria+ecto, data=md, random=~1|name))

```

And the results are very similar as with richness: Eimeria is the only significant covariate.

Let's generate the figures: first Figure 1a.

```{r}
multiple <- c("Anisha","Ari","Benja","Chewbacca","Cleopatra","Lasse","Laura","Lita","Napolean","Rachootin","Raleva") #mouse lemurs with multiple samples which were positive and negative for eimeria
md_eim <- subset(md, name %in% multiple)
md_noeim <- subset(md, !name %in% multiple)

ggplot()+
  geom_jitter(data=md_eim, aes(eimeria,sobs,color=eimeria),position=position_jitter(0.1))+
  stat_summary(data=md_eim, aes(eimeria,sobs,color=eimeria),fun.y=median, geom="crossbar", ymax=-100, ymin=-100)+
  geom_jitter(data=md_noeim, aes(eimeria,sobs),position=position_jitter(0.1),colour="grey")+
  stat_summary(data=md, aes(eimeria,sobs),fun.y=median, geom="crossbar", ymax=-100, ymin=-100,colour="grey")+
  labs(x="",y="Richness")+
  theme(axis.text.x=element_blank(),
        axis.text.y=element_text(colour="black"),
        panel.background=element_rect(fill="white", colour="black"),
        strip.background=element_blank())
```
And then Figure 1b
```{r}
ggplot()+
  geom_jitter(data=md_eim, aes(eimeria,invsimpson,color=eimeria),position=position_jitter(0.1))+
  stat_summary(data=md_eim, aes(eimeria,invsimpson,color=eimeria),fun.y=median, geom="crossbar", ymax=-100, ymin=-100)+
  geom_jitter(data=md_noeim, aes(eimeria,invsimpson),position=position_jitter(0.1),colour="grey")+
  stat_summary(data=md, aes(eimeria,invsimpson),fun.y=median, geom="crossbar", ymax=-100, ymin=-100,colour="grey")+
  labs(x="",y="Diversity")+
  theme(axis.text.x=element_blank(),
        axis.text.y=element_text(colour="black"),
        panel.background=element_rect(fill="white", colour="black"),
        strip.background=element_blank())

```

Then we move to dissimilarity matrices which were presented in the Appendix of the manuscript.

First we load them, and do some housekeeping.

```{r}
dist <- read.table(file = "yueclayton.dist", 
                   skip = 1, row.names = 1)

dist <- dist[row.names(dist) %in% md$group,]
rownames(dist) <- gsub("^(.*)_16S", "\\1", rownames(dist))
o <- order(rownames(dist))
dist <- dist[o, o]
stopifnot(sum(rownames(dist) != rownames(md)) == 0)
yueclay <- dist
thetayc <- as.dist(dist)

dist <- read.table(file = "jaccard.dist", 
                   skip = 1, row.names = 1)
dist <- dist[row.names(dist) %in% md$group,]
rownames(dist) <- gsub("^(.*)_16S", "\\1", rownames(dist))
o <- order(rownames(dist))
dist <- dist[o, o]
stopifnot(sum(rownames(dist) != rownames(md)) == 0)
jacc <- dist
jclass <- as.dist(dist)

dist <- read.table(file = "braycurtis.dist", 
                   skip = 1, row.names = 1)
dist <- dist[row.names(dist) %in% md$group,]
rownames(dist) <- gsub("^(.*)_16S", "\\1", rownames(dist))
o <- order(rownames(dist))
dist <- dist[o, o]
stopifnot(sum(rownames(dist) != rownames(md)) == 0)
bray <- dist
braycurtis <- as.dist(dist)


dist <- read.table(file = "ochiai.dist", 
                   skip = 1, row.names = 1)
dist <- dist[row.names(dist) %in% md$group,]
rownames(dist) <- gsub("^(.*)_16S", "\\1", rownames(dist))
o <- order(rownames(dist))
dist <- dist[o, o]
mdo = md [-30,] 
stopifnot(sum(rownames(dist) != rownames(mdo)) == 0)
och <- dist
ochiai <- as.dist(dist)

md$week <- as.factor(md$week)
```

Next, we will do the permutational manovas.

```{r}
adonis(thetayc ~ md$age+md$cond+md$week+md$aggr+md$sex+md$site+md$nematode+md$hnana+md$hdiminuta+md$eimeria+md$ecto, strata = md$name, perm = 1000) 

adonis(jclass ~ md$age+md$sex+md$site+md$nematode+md$hnana+md$week+md$hdiminuta+md$eimeria+md$ecto, strata=md$name, perm = 1000)

adonis(braycurtis ~ md$seqbatch+md$age+md$sex+md$site+md$week+md$nematode+md$hnana+md$hdiminuta+md$eimeria+md$ecto, strata=md$name,perm = 1000)

adonis(ochiai ~ mdo$week+mdo$age+mdo$sex+mdo$site+mdo$nematode+mdo$hnana+mdo$hdiminuta+mdo$eimeria+mdo$ecto,strata=mdo$name,perm = 1000) 

```

...and accompanying Figure SXX in Appendix 2. First Yue-Clayton:

```{r}
mod_yc <- metaMDS(thetayc)
MDS1 <- as.numeric(mod_yc$points[,1])
MDS2 <- as.numeric(mod_yc$points[,2])
MDS <- cbind(MDS1, MDS2)
row.names(MDS) <- row.names(md$points)
MDS_yc <- data.frame(MDS)
MDS_yc$ecto <- md$ecto
MDS_yc$comb <- paste(md$ecto,md$hdiminuta, sep="_")
MDS_yc_group <- as.factor(MDS_yc$ecto)

find_hull <- function(MDS_yc) MDS_yc[chull(MDS_yc$MDS1,MDS_yc$MDS2), ]
hulls <- ddply(MDS_yc, "MDS_yc_group", find_hull)
hulls$ecto <- as.factor(hulls$ecto)

ggplot(NULL, aes())+
  geom_polygon(data=hulls, aes(MDS1,MDS2,fill=factor(ecto)), alpha=0.1)+
  geom_point(data=MDS_yc, aes(MDS1,MDS2,shape=factor(comb)), alpha=1, size=4)+
  scale_shape_manual(values=c(4,2,16,17))+
  labs(title="Yue-Clayton \nEctoparasites", x="Axis 1",y="Axis 2")+
  theme(axis.text=element_text(colour="black"),
        legend.title=element_blank(),
        panel.background=element_rect(fill="white", colour="black"))
```

Then Jaccard.

```{r}

mod_j <- metaMDS(jclass)
MDS1 <- as.numeric(mod_j$points[,1])
MDS2 <- as.numeric(mod_j$points[,2])
MDS <- cbind(MDS1, MDS2)
row.names(MDS) <- row.names(md$points)
MDS_j <- data.frame(MDS)
MDS_j$hdiminuta <- md$hdiminuta
MDS_j$comb <- paste(md$ecto, md$hdiminuta, sep="_")
MDS_j_group <- as.factor(MDS_j$hdiminuta)

find_hull <- function(MDS_j) MDS_j[chull(MDS_j$MDS1,MDS_j$MDS2), ]
hulls <- ddply(MDS_j, "MDS_j_group", find_hull)
hulls$hdiminuta <- as.factor(hulls$hdiminuta)

ggplot(NULL, aes())+
  geom_polygon(data=hulls, aes(MDS1,MDS2,fill=factor(hdiminuta)), alpha=0.1)+
  geom_point(data=MDS_j, aes(MDS1,MDS2,shape=factor(comb)), alpha=1, size=4)+
  scale_shape_manual(values=c(4,2,16,17))+
  labs(title="Jaccard \nH.diminuta", x="Axis 1",y="Axis 2")+
  theme(axis.text=element_text(colour="black"),
        legend.title=element_blank(),
        panel.background=element_rect(fill="white", colour="black"))
```

Bray-Curtis:

```{r}

mod_bc <- metaMDS(braycurtis)
MDS1 <- as.numeric(mod_bc$points[,1])
MDS2 <- as.numeric(mod_bc$points[,2])
MDS <- cbind(MDS1, MDS2)
row.names(MDS) <- row.names(md$points)
MDS_bc <- data.frame(MDS)
MDS_bc$factor <- paste(md$ecto,md$hdiminuta, sep="_")
MDS_bc_group <- as.factor(MDS_bc$factor)

find_hull <- function(MDS_bc) MDS_bc[chull(MDS_bc$MDS1,MDS_bc$MDS2), ]
hulls <- ddply(MDS_bc, "MDS_bc_group", find_hull)
hulls$factor <- as.factor(hulls$factor)

ggplot(NULL, aes(fill=factor(factor), shape=factor(factor)))+
  geom_polygon(data=hulls, aes(MDS1,MDS2),alpha=0.1)+
  geom_point(data=MDS_bc, aes(MDS1,MDS2), alpha=1, size=4)+
  scale_shape_manual(values=c(4,2,16,17))+
  labs(title="Bray-Curtis \nEctoparasites - H.diminuta", x="Axis 1",y="Axis 2")+
  theme(axis.text=element_text(colour="black"),
        legend.title=element_blank(),
        panel.background=element_rect(fill="white", colour="black"))
```

No need for Ochiai as there were no significant variables.

Next we go to group dispersion, which takes a bit of work because there is no ready-made replication within the #function. Thus we need take one sample per lemur and iterate this. First we create the needed iteration.

```{r}
names <- levels(droplevels(md$name))
name_nr <- length(names)
rows_with <- numeric(name_nr)
```

We define function resample() based on the documentation on sample()

```{r}
resample <- function(x, ...) x[sample.int(length(x), ...)]
```

Then we create null vectors and resample each distance matrix (5 parasites X 4 metrics):

```{r}
# creating vectors
nema_yc_iter=NULL
nema_bc_iter=NULL
nema_j_iter=NULL
nema_o_iter=NULL
hnana_yc_iter=NULL
hnana_bc_iter=NULL
hnana_j_iter=NULL
hnana_o_iter=NULL
hdimi_yc_iter=NULL
hdimi_bc_iter=NULL
hdimi_j_iter=NULL
hdimi_o_iter=NULL
eim_yc_iter=NULL
eim_bc_iter=NULL
eim_j_iter=NULL
eim_o_iter=NULL
ecto_yc_iter=NULL
ecto_bc_iter=NULL
ecto_j_iter=NULL
ecto_o_iter=NULL

#resampling for each distance matrix

for (i in 1:100){
  
  for(j in 1:name_nr){
    rows_with[j] <- resample(which(md$name == names[j]),1)
  }
  
  
  rows_with <- sort(rows_with)
  mdr <- md[rows_with,]
  
  thetayc2 <- as.dist(yueclay[rows_with,rows_with])
  betadisper_ycnem <- betadisper(thetayc2,mdr$nematode)
  anova(betadisper_ycnem)
  perm <- permutest(betadisper_ycnem, pairwise=TRUE)
  nema_yc_iter[i] = (perm$pairwise$observed[1]<0.05)
  
  betadisper_ychn <- betadisper(thetayc2,mdr$hnana)
  anova(betadisper_ychn)
  perm <- permutest(betadisper_ychn, pairwise=TRUE)
  hnana_yc_iter[i] = (perm$pairwise$observed[1]<0.05)
  
  betadisper_ychd <- betadisper(thetayc2,mdr$hdiminuta)
  anova(betadisper_ychd)
  perm <- permutest(betadisper_ychd, pairwise=TRUE)
  hdimi_yc_iter[i] = (perm$pairwise$observed[1]<0.05)
  
  betadisper_ycei <- betadisper(thetayc2,mdr$eimeria)
  anova(betadisper_ycei)
  perm <- permutest(betadisper_ycei, pairwise=TRUE)
  eim_yc_iter[i] = (perm$pairwise$observed[1]<0.05)
  
  betadisper_ycec <- betadisper(thetayc2,mdr$ecto)
  anova(betadisper_ycec)
  perm <- permutest(betadisper_ycec, pairwise=TRUE)
  ecto_yc_iter[i] = (perm$pairwise$observed[1]<0.05)
  
}

for (i in 1:100){
  
  for(j in 1:name_nr){
    rows_with[j] <- resample(which(md$name == names[j]),1)
  }
  
  
  rows_with <- sort(rows_with)
  mdr <- md[rows_with,]
  
  thetayc2 <- as.dist(bray[rows_with,rows_with])
  betadisper_bcnem <- betadisper(thetayc2,mdr$nematode)
  anova(betadisper_bcnem)
  perm <- permutest(betadisper_bcnem, pairwise=TRUE)
  nema_bc_iter[i] = (perm$pairwise$observed[1]<0.05)
  
  betadisper_bchn <- betadisper(thetayc2,mdr$hnana)
  anova(betadisper_bchn)
  perm <- permutest(betadisper_bchn, pairwise=TRUE)
  hnana_bc_iter[i] = (perm$pairwise$observed[1]<0.05)
  
  betadisper_bchd <- betadisper(thetayc2,mdr$hdiminuta)
  anova(betadisper_bchd)
  perm <- permutest(betadisper_bchd, pairwise=TRUE)
  hdimi_bc_iter[i] = (perm$pairwise$observed[1]<0.05)
  
  betadisper_bcei <- betadisper(thetayc2,mdr$eimeria)
  anova(betadisper_bcei)
  perm <- permutest(betadisper_bcei, pairwise=TRUE)
  eim_bc_iter[i] = (perm$pairwise$observed[1]<0.05)
  
  betadisper_bcec <- betadisper(thetayc2,mdr$ecto)
  anova(betadisper_bcec)
  perm <- permutest(betadisper_bcec, pairwise=TRUE)
  ecto_bc_iter[i] = (perm$pairwise$observed[1]<0.05)
  
}
for (i in 1:100){
  
  for(j in 1:name_nr){
    rows_with[j] <- resample(which(md$name == names[j]),1)
  }
  
  rows_with <- sort(rows_with)
  mdr <- md[rows_with,]
  
  thetayc2 <- as.dist(jacc[rows_with,rows_with])
  betadisper_jnem <- betadisper(thetayc2,mdr$nematode)
  anova(betadisper_jnem)
  perm <- permutest(betadisper_jnem, pairwise=TRUE)
  nema_j_iter[i] = (perm$pairwise$observed[1]<0.05)
  
  betadisper_jhn <- betadisper(thetayc2,mdr$hnana)
  anova(betadisper_jhn)
  perm <- permutest(betadisper_jhn, pairwise=TRUE)
  hnana_j_iter[i] = (perm$pairwise$observed[1]<0.05)
  
  betadisper_jhd <- betadisper(thetayc2,mdr$hdiminuta)
  anova(betadisper_jhd)
  perm <- permutest(betadisper_jhd, pairwise=TRUE)
  hdimi_j_iter[i] = (perm$pairwise$observed[1]<0.05)
  
  betadisper_jei <- betadisper(thetayc2,mdr$eimeria)
  anova(betadisper_jei)
  perm <- permutest(betadisper_jei, pairwise=TRUE)
  eim_j_iter[i] = (perm$pairwise$observed[1]<0.05)
  
  betadisper_jec <- betadisper(thetayc2,mdr$ecto)
  anova(betadisper_jec)
  perm <- permutest(betadisper_jec, pairwise=TRUE)
  ecto_j_iter[i] = (perm$pairwise$observed[1]<0.05)
}


for (i in 1:100){
  
  for(j in 1:name_nr){
    rows_with[j] <- resample(which(mdo$name == names[j]),1)
  }
  
  rows_with <- sort(rows_with)
  mdr <- mdo[rows_with,]
  
  thetayc2 <- as.dist(och[rows_with,rows_with])
  betadisper_onem <- betadisper(thetayc2,mdr$nematode)
  anova(betadisper_onem)
  perm <- permutest(betadisper_onem, pairwise=TRUE)
  nema_o_iter[i] = (perm$pairwise$observed[1]<0.05)
  
  betadisper_ohn <- betadisper(thetayc2,mdr$hnana)
  anova(betadisper_ohn)
  perm <- permutest(betadisper_ohn, pairwise=TRUE)
  hnana_o_iter[i] = (perm$pairwise$observed[1]<0.05)
  
  betadisper_ohd <- betadisper(thetayc2,mdr$hdiminuta)
  anova(betadisper_ohd)
  perm <- permutest(betadisper_ohd, pairwise=TRUE)
  hdimi_o_iter[i] = (perm$pairwise$observed[1]<0.05)
  
  betadisper_oei <- betadisper(thetayc2,mdr$eimeria)
  anova(betadisper_oei)
  perm <- permutest(betadisper_oei, pairwise=TRUE)
  eim_o_iter[i] = (perm$pairwise$observed[1]<0.05)
  
  betadisper_oec <- betadisper(thetayc2,mdr$ecto)
  anova(betadisper_oec)
  perm <- permutest(betadisper_oec, pairwise=TRUE)
  ecto_o_iter[i] = (perm$pairwise$observed[1]<0.05)
  
}
```

Then we calculate the positive samples:

```{r, eval=FALSE}
sum(hnana_bc_iter)
sum(hnana_j_iter)
sum(hnana_o_iter)
sum(hnana_yc_iter)

sum(hdimi_bc_iter)
sum(hdimi_j_iter)
sum(hdimi_o_iter)
sum(hdimi_yc_iter)

sum(nema_bc_iter)
sum(nema_j_iter)
sum(nema_o_iter)
sum(nema_yc_iter)

sum(eim_bc_iter)
sum(eim_j_iter)
sum(eim_o_iter)
sum(eim_yc_iter)

sum(ecto_bc_iter)
sum(ecto_j_iter)
sum(ecto_o_iter)
sum(ecto_yc_iter)
```

These were the results I got from my definite run:

Number of significant group dispersions with 100 iterations:
	 	 	


| | Jaccard | Ochiai | Bray-Curtis | Yue-Clayton |
|------:|:-----|:---------|:------|:------|
|H. diminuta|0|0|0|0|
|H. nana|0|0|0|0|
|Ectos|1|22|0|11|
|Eimeria|8|4|0|0|
|Nematode|0|11|4|10|	

Last - and least - the effect of new infections on similarities

```{r}
simi_par <- read.csv(file="interaction_similarity.csv", sep=";", dec=",")
simi_par$hnana <- as.factor(simi_par$hnana)
simi_par$hdiminuta <- as.factor(simi_par$hdiminuta)
simi_par$nematode <- as.factor(simi_par$nematode)
simi_par$ecto <- as.factor(simi_par$ecto)
simi_par$eimeria <- as.factor(simi_par$eimeria)

anova(lme(simil_jac ~ nematode+hnana+hdiminuta+eimeria+ecto+total+days, data=simi_par, random=~1|name))

anova(lme(simil_the ~ nematode+hnana+hdiminuta+eimeria+ecto+total+days, data=simi_par, random=~1|name))
```